Revisiting the Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation using complex Empirical Orthogonal Functions

### 2021-08-13 analysis/paper/shceof.Rmd
Cambios en figuras

width_column <- 3.6

@@

```{r eof-naive, fig.cap = "(ref:eof-naive-cap)", fig.width=2.5, fig.height=3.2}
eof_naive[, eof[[1]]$right, by = lev] %>%
  .[lev == 50] %>% 
  .[ , hgt := hgt/sd(hgt)] %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = hgt, fill = ..level..),
                    breaks = AnchorBreaks(0, 1, exclude = 0)) +
  geom_contour_tanaka(aes(z = hgt), breaks = AnchorBreaks(0, 1, exclude = 0), 
                      range = c(0.01, 0.2)) +

@@

\textcolor{red}{Fig. 6. Te quedan muy pegados los subplots de la columna izquierda con los de la derecha. Recomendaría aumentar el espacio intermedio y repetir el eje y en las figuras de la derecha. Además podrías agregar el y-label (Level) a la izquierda.}

@@

\textcolor{red}{Tabla 2. Mismo comentario que antes con lo del background color (no dice mucho si no se conoce la escala), quizás podes destacar en negritas según la significancia. ¿Está bien el intervalo de confianza de PSA2 vs cEOF2 Im? Hay algo raro ahí}

@@

```{r enso_cor}
enso_cor <-  ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm() %>%
  .[, hgt := scale(hgt), by = .(cEOF, part)] %>%
  enso[., on = "time"] %>%
  na.omit() %>%
  .[, correlate(hgt, oni), by = .(cEOF, part)]
```

@@

    geom_contour_fill(data = sst_gdata,
                      aes(z = estimate, fill = ..level..), 
                      breaks = breaks_sst) +
    geom_contour_pval(data = sst_gdata, aes(z = p_val)) +
    scale_fill_divergent_discretised("SST (K)", 
                                     labels = one_less,
                                     guide = guide_colorsteps_bottom(width = 15,
                                                                     show.limits = FALSE, ticks = TRUE,
                                                                     order = 1)) +
    scale_x_longitude() +
    scale_y_latitude() +
    scale_mag(guide = "none") +
    facet_grid(term ~ var) +
    # coord_quickmap(ylim = c(NA, 10)) +
    coord_sf(ylim = c(-90, 0),
             crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +over") + 
    theme(legend.box = "horizontal", legend.text = element_text(size = rel(0.7))) +
    tag_facets("cr", position = "bl") 
}

@@

```{r sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(which_cEOF = "cEOF2", which_lev = 200)

@@

Since the Real cEOF2 represents just a different phase of the same wave train, it would be expected to show a similar forcing pattern to the Imaginary cEOF2 with a slight translation of its location.
However, Figure \@ref(fig:sst-psi-2).a1 and b1 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics.
The correlation between the Real cEOF2 and ENSO is also not significant (`r enso_cor[cEOF == "cEOF2" & part == "Real"]$text`).
\textcolor{red}{Usaría notación científica por la cantidad de 0 que tiene}.
This lack of tropical signal suggest that the Real cEOF2 and the Imaginary cEOF2 have very different sources of variability.

@@

\textcolor{red}{Fig. 12 Revisa cómo te quedo abajo de la figura la indicación de qué son cada uno de los puntos (te quedaron dos circulos y una línea extra que no va)}

@@

\textcolor{red}{Fig. 13 Tene en cuenta los mismos comentarios que en el figura 11 y revisa también la barra de colores para SST que quedo rara}

@@

    scale_x_longitude(ticks = 15) +
    scale_y_latitude(ticks = 15) +
    ggh4x::facet_nested(term ~ cEOF) +
    axis_labs_smol +
    tag_facets("cr") +
    coord_sf(ylim = continents[[which_continent]]$lat,
             xlim = continents[[which_continent]]$lon,
             crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +over")

@@

(ref:pp-oceania-cap) Regression of SON mean precipitation anomalies in Oceania (mm per day, shaded) and (column a) cEOF1 the (row 1) Imaginary and (column 1) Real phase. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate. In black is the West Australia region, defined as the box between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))`.

\textcolor{red}{Revisa el caption de la figura 14 que equedo un estilo distinto que los anteriores. Te falto aclarar también qué es la caja (¿la inventaste vos o habías utilizado la de algún otro trabajo?). Fijate también de aumentar la separación entre la columna izquierda y derecha para que no se superpongan los labels. También fijate que esta barra de colores no te quedo con el borde externo como el resto. }

```{r pp-oceania, fig.cap = "(ref:pp-oceania-cap)", fig.width=width_column, fig.height=width_column*.8}
plot_pp(regr_pp, c("cEOF1", "cEOF2"), which_continent =  "oceania") +
  geom_rect(data = box_australia, aes(xmin = min(lon), xmax = max(lon),
                                      ymin = min(lat), ymax = max(lat)),
            fill = NA, colour = "gray20") 
```

The precipitation anomalies in Oceania associated with cEOF1 are weak and not statistically significant (Figure \@ref(fig:pp-oceania) column a) whereas cEOF2 is associated with large and widespread anomalies. The Real cEOF2 is associated with positive seasonal anomalies in Northern and Eastern Australia, and negative ones over New Zealand, while precipitation anomalies associated with Imaginary cEOF2 are negative over all Eastern Australia. As it could be expected, these negative anomalies are similar to the springtime precipitation anomalies observed in relation with the Asymmetric SAM [@campitelli2021] and ENSO [@cai2011].
\textcolor{red}{Acá te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

@@

\textcolor{red}{Revisa lo mismo que para la otra figura del tipo. }

@@

Figure \@ref(fig:pp-america) shows regression maps of seasonal precipitation with each cEOF in South America. cEOF1 is associated with negative regressions coefficients for precipitation in Southern Brazil and Paraguay, although these are not statistically significant (Fig. \@ref(fig:pp-america) column a). The strongest precipitation regression coefccients are the ones associated with the Imaginary cEOF2. The positive regression coefficients on Southeastern South America (SESA) and Central Chile, and negative ones over South Atlantic Convergence Zone is a well-known springtime precipitation signature of ENSO [@cai2020a] and it is also similar to the precipitation anomalies associated with the A-SAM [@campitelli2021].
\textcolor{red}{Lo mismo que te puse antes. Te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

@@

(ref:pp-africa-cap) Same as Figure \@ref(fig:pp-america) but for South of Africa.

```{r pp-africa, fig.cap = "(ref:pp-africa-cap)"}
# plot_pp(regr_pp, c("cEOF1", "cEOF2"), which_continent = "africa")
```

>>>>

width_column <- 3.3

@@

```{r eof-naive, fig.cap = "(ref:eof-naive-cap)", fig.width=width_column, fig.height=width_column*1.2}
eof_naive[, eof[[1]]$right, by = lev] %>%
  .[lev == 50] %>% 
  .[ , hgt := hgt/sd(hgt)] %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = hgt, fill = ..level..),
                    breaks = AnchorBreaks(0, 1, exclude = 0)) +
  geom_contour_tanaka(aes(z = hgt), breaks = AnchorBreaks(0, 1, exclude = 0), 
                      range = c(0.01, 0.2)) +

@@

@@

@@

```{r enso_cor}
enso_cor <-  ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm() %>%
  .[, hgt := scale(hgt), by = .(cEOF, part)] %>%
  enso[., on = "time"] %>%
  na.omit() %>%
  .[, correlate(hgt, oni), by = .(cEOF, part)]

@@

    geom_contour_fill(data = sst_gdata,
                      aes(z = estimate, fill = ..level..), 
                      breaks = breaks_sst) +
    geom_contour_pval(data = sst_gdata, aes(z = p_val)) +
    scale_fill_divergent_discretised("SST (K)", 
                                     labels = one_less,
                                     guide = guide_colorsteps_bottom(width = 15,
                                                                     show.limits = FALSE, ticks = TRUE,
                                                                     order = 1)) +
    scale_x_continuous(NULL, expand = c(0, 0), 
                       breaks = seq(-180, 180-60, by = 60),
                       labels = LonLabel) +

    scale_y_continuous(NULL, breaks  = seq(-90, 0, by = 30), expand = c(0, 0)) +
    scale_mag(guide = "none") +
    facet_grid(term ~ var) +
    # coord_quickmap(ylim = c(NA, 10)) +
    coord_sf(ylim = c(-89, 0), expand = FALSE,
             crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +lon_wrap=180 +over") + 
    theme(legend.box = "horizontal",
          legend.text = element_text(size = rel(0.7)), 
          legend.box.spacing = grid::unit(0, "lines") ) +
    tag_facets("cr", position = "bl") 
}

@@

```{r sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 7, fig.height=4, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(which_cEOF = "cEOF2", which_lev = 200) 
```

@@

Since the Real cEOF2 represents just a different phase of the same wave train, it would be expected to show a similar forcing pattern to the Imaginary cEOF2 with a slight translation of its location.
However, Figure \@ref(fig:sst-psi-2).a1 and b1 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics.
The correlation between the Real cEOF2 and ENSO is also not significant (`r enso_cor[cEOF == "cEOF2" & part == "Real"]$text`).
This lack of tropical signal suggest that the Real cEOF2 and the Imaginary cEOF2 have very different sources of variability.

@@

@@

@@

    scale_x_continuous("", expand = c(0, 0), breaks = seq(0, 360, by = 15), 
                       labels = LonLabel) +
    scale_y_latitude(ticks = 15) +
    ggh4x::facet_nested(term ~ cEOF) +
    axis_labs_smol +
    tag_facets("rc") +
    coord_sf(ylim = continents[[which_continent]]$lat,
             xlim = continents[[which_continent]]$lon,
             crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +over") +
    theme(panel.spacing = grid::unit(1, "lines"))

@@

(ref:pp-oceania-cap) Regression of SON mean precipitation anomalies in Oceania (mm per day, shaded) with the (row a) real and (row b) imaginary pars of the (column 1) cEOF1 and (column 2) cEOF2. For the 1979 -- 2019. Areas marked with dots have p-values smaller than 0.05 adjusted for False Detection Rate. In black is the West Australia region, defined as the box between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))`.

```{r pp-oceania, fig.cap = "(ref:pp-oceania-cap)", fig.width=width_column, fig.height=width_column*.8}
plot_pp(regr_pp, c("cEOF1", "cEOF2"), which_continent =  "oceania") +
  geom_rect(data = box_australia, aes(xmin = min(lon), xmax = max(lon),
                                      ymin = min(lat), ymax = max(lat)),
            fill = NA, colour = "gray20")

The precipitation anomalies in Oceania associated with cEOF1 are weak and not statistically significant (Figure \@ref(fig:pp-oceania) column 1) whereas cEOF2 is associated with large and widespread anomalies. The Real cEOF2 is associated with positive seasonal anomalies in Northern and Eastern Australia, and negative ones over New Zealand, while precipitation anomalies associated with Imaginary cEOF2 are negative over all Eastern Australia. As it could be expected, these negative anomalies are similar to the springtime precipitation anomalies observed in relation with the Asymmetric SAM [@campitelli2021] and ENSO [@cai2011].
\textcolor{red}{Acá te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

@@

@@

Figure \@ref(fig:pp-america) shows regression maps of seasonal precipitation with each cEOF in South America. cEOF1 is associated with negative regressions coefficients for precipitation in Southern Brazil and Paraguay, although these are not statistically significant (Fig. \@ref(fig:pp-america) column 1). The strongest precipitation regression coefficients are the ones associated with the Imaginary cEOF2. The positive regression coefficients on Southeastern South America (SESA) and Central Chile, and negative ones over South Atlantic Convergence Zone is a well-known springtime precipitation signature of ENSO [@cai2020a] and it is also similar to the precipitation anomalies associated with the A-SAM [@campitelli2021].
\textcolor{red}{Lo mismo que te puse antes. Te faltaría  sacar un poco más el jugo asociando con la figura 11. Además, ¿revisaste si podes también usar algún resultado asociado a los PSA? }

@@
